---
title: "hw04CarolineSeip"
author: "Caroline Seip"
date: "October 3, 2018"
output: github_document
---
#Table of Contents
- Data reshaping: Activity #2
- 

#Data reshaping: Activity #2

Make a tibble with one row per year and columns for life expectancy for two or more countries.
- Use knitr::kable() to make this table look pretty in your rendered homework.
- Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

First let's load the dataset, tidyverse and knitr:

```{r}
#Load gapminder dataset
library(gapminder)
#Load tidyverse to use dplyr and ggplot
library(tidyverse)
#Load knitr to use the kable function
library(knitr)
```

Now let's make a formatted table of life expectancy for each year in Australia and New Zealand:

```{r}
#Use the gapminder dataset
gapminder %>%
  #Filter so that we are only using Oceania (Australia and New Zealand)
  filter(continent=="Oceania") %>% 
  #Select the columns country, year and life expectancy
  select(country, year, lifeExp) %>% 
  #Make a table, rename column headings
  kable(col.names = c("Country", "Year", "Life expectancy (years)"))
```


