---
title: "hw04CarolineSeip"
author: "Caroline Seip"
date: "October 3, 2018"
output: github_document
---
#Table of Contents
- Data reshaping: Activity #2
- 

#Data reshaping: Activity #2

**The prompt I have chosen to explore is Activity 2:**

- Make a tibble with one row per year and columns for life expectancy for two or more countries
- Use knitr::kable() to make this table look pretty in your rendered homework
- Take advantage of this new data shape to scatterplot life expectancy for one country against that of another

First let's load the dataset, tidyverse and knitr:

```{r}
#Load gapminder dataset
library(gapminder)
#Load tidyverse to use dplyr and ggplot
library(tidyverse)
#Load knitr to use the kable function
library(knitr)
```

Now let's make a formatted table of life expectancy for each year in Australia and New Zealand:

```{r}
#Create an object 'OceaniaLifeExp' to store the data in
OceaniaLifeExp <- 
  #Use the gapminder dataset
  gapminder %>%
  #Filter so that we are only using Oceania (Australia and New Zealand)
  filter(continent=="Oceania") %>% 
  #Select the columns country, year and life expectancy
  select(country, year, lifeExp)
  #Make a tidy table using the 'kable' function, rename column headings
  kable(OceaniaLifeExp, col.names = c("Country", "Year", "Life expectancy (years)"))
```

This is great, but we want to organize the data now so that we only have one row for each year, so we are making the data 'untidy':

```{r}
#Use the 'spread' command to untidy the data, specify the key (column titles) and value (column values)
spread(OceaniaLifeExp, key = "country", value = "lifeExp") %>% 
  #Make a pretty table using the 'kable' command, and specify column names
  kable(col.names = c("Year", "Australia", "New Zealand"))
```

Now let's make a scatterplot to compare life expectancy in New Zealand and Australia:

```{r}
#Use the previously created object 'OceaniaLifeExp'
OceaniaLifeExp %>% 
  #Use ggplot and specify the aesthetics (x and y), 'year' and 'lifeExp'
  ggplot(aes(year, lifeExp)) +
  #Add a scatterplot, colour the points by country
  geom_point(aes(colour= country)) +
  #Add a title
  ggtitle("Life Expectancies in Oceania 1952-2007") +
  #Add x and y labels
  xlab("Year") +
  ylab("Life expectancy (years)")
```


